---
title: "Assigment for Data Wrangling and Vizualization Course"
subtitle: "Plotting Assigment "
author: "Katrin Grings"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    theme: cosmo
    toc: yes
    toc_float:
      collapsed: true
---

```{r, include=FALSE}

knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE)

```

# Dependencies

```{r}

library(dplyr)
library(tidyr)
library(readr)
library(janitor)
library(stringr)
library(ggplot2)
library(ggridges)
library(ggtext)
library(ggdist)
library(glue)
library(patchwork)
library(showtext)
library(devtools)
devtools::install_github("schochastics/Rokemon")
library(Rokemon)
library(extrafont)
```

# Get data

```{r}

#Load Pokemon dataset for bad plot
pokemon_dataset <- read_delim("../data/pokemon_dataset.csv", 
    delim = ",", escape_double = FALSE, col_names = TRUE, show_col_types = FALSE,
    trim_ws = TRUE)

```
# Data Wrangling  

## Data Wrangling Good Plot

```{r}
#Reorder weight to descending mean
data_good_plot <- chickwts %>%
  mutate(feed = reorder(feed, -weight, FUN = mean))

#Mean weight overall
mean_weight_all <- mean(chickwts$weight, na.rm = TRUE)

#New data frame for mean and number of chickens per feed type
mean_weight_feed <- chickwts %>%
  group_by(feed) %>%
  summarise(
    weight = mean(weight, na.rm = TRUE),
    count = n()
  )

# Filter for one feed type to create a custom legend plot
df_for_legend <- chickwts %>% 
  filter(feed == "sunflower")

```

## Data Wrangling Bad Plot

```{r}

# Filter Pokemon from the first Generation and handle missing secondary types
data_bad_plot <- pokemon_dataset %>%
  filter(generation == "gen 1") %>%
  mutate(
    secondary_type = replace_na(secondary_type, "no") # Replace missing secondary types with "no"
  )

```

# Plots

## Good Plot 

### Plot Specifications

```{r}

font_add_google("Fira Sans", "fira_sans")

#Define color and font used in the plot
gp_color <- "grey97"
gp_font_family <- "fira_sans"

#Plot subtitle
plot_subtitle = glue("An experiment was conducted to measure and compare the effectiveness of various feed supplements on the growth rate of chickens. Newly hatched chicks were randomly allocated into six groups, and each group was given a different feed supplement. Their weights in grams after six weeks are given along with feed types.
")


```

### Main plot

```{r}

good_plot_main <- data_good_plot %>%
  ggplot(aes(feed, weight)) +
  
# Distribution of weights
  stat_halfeye(fill_type = "segments", alpha = 0.8, width = 0.8) +
  
# Bars below distribution
  stat_interval() +
  
# Mean of each feed type as a point
  stat_summary(geom = "point", fun = mean, color = "yellow", size = 2) +
    
  geom_point(alpha = 0.5)+
  
# Mean weight line
  geom_hline(yintercept = mean_weight_all, col = "grey30", lty = "dashed") +
  annotate("text", x = 7, y = mean_weight_all, label = "Mean weight",
           family = gp_font_family, size = 3, hjust = 0) +
  
# Annotation for number of chicks per feed type
  geom_text(data = mean_weight_feed,
            aes(x = feed, y = 75, label = sprintf("(%d)", count)),
            family = gp_font_family, size = 3, hjust = 1) +
  annotate("text", x = 7, y = 75, label = "(n Chickens)",
           family = gp_font_family, size = 3, hjust = 0.5) +
  
# Scale and label adjustments
  scale_x_discrete(labels = toupper, expand = c(0.1, 0)) +
  scale_y_continuous(breaks = seq(100, 400, 50)) +
  scale_color_manual(values = MetBrewer::met.brewer("Klimt")) +
  coord_flip(ylim = c(50, 450), clip = "off") +
  guides(col = "none") +
  
# Labels
  labs(
    title = toupper("Chicken Weight per Feed Type"),
    subtitle = plot_subtitle,
    caption = "Axis capped at 500 grams <br>
    Data: McNeil, D. R. (1977). <br>
    Retrieved from Dataset chickwts.",
    family = gp_font_family,
    x = NULL,
    y = "Chicken Weights"
  ) +
  
# Aesthetic adjustments 
  theme_minimal(base_family = gp_font_family) +
  theme(
    plot.background = element_rect(color = NA, fill = gp_color),
    panel.grid = element_blank(),
    panel.grid.major.x = element_line(linewidth = 0.1, color = "grey75"),
    plot.title = element_text(family = gp_font_family),
    plot.title.position = "plot",
    plot.subtitle = element_textbox_simple(margin = margin(t = 4, b = 16), size = 10),
    plot.caption = element_textbox_simple(margin = margin(t = 12), size = 7),
    plot.caption.position = "plot",
    axis.text.y = element_text(hjust = 0.3, margin = margin(r = 0), family = gp_font_family),
    panel.spacing = unit(1, "lines"),
    plot.margin = margin(4, 4, 4, 4)
  )

```


### Legend Plot

```{r}

good_plot_legend <- df_for_legend %>%
  ggplot(aes(feed, weight)) +
  
# Plot elements
  stat_halfeye(fill_type = "segments", alpha = 0.3) +
  stat_interval() +
  stat_summary(geom = "point", fun = median, color = "yellow", size = 2) +
  
# Annotations with descriptions
  annotate(
    "richtext",
    x = c(0.5, 0.8, 0.8, 1.6, 2),
    y = c(320, 250, 400, 285, 225),
    label = c(
      "50 % of weight <br>fall within this range", 
      "95 % of weight", "80 % of weight", 
      "Mean", "Distribution<br>of weights"
    ),
    fill = NA, label.size = NA, family = gp_font_family, size = 2, vjust = 1
  ) +
  
# Arrows for description annotations
  geom_curve(
    data = data.frame(
      x = c(0.50, 0.75, 0.75, 1.4, 2),
      xend = c(0.90, 0.90, 0.90, 1.1, 1.8), 
      y = c(320, 250, 400, 300, 250),
      yend = c(330, 275, 370, 325, 325)
    ),
    aes(x = x, xend = xend, y = y, yend = yend),
    stat = "unique", curvature = 0.2, size = 0.2, color = "grey12",
    arrow = arrow(angle = 20, length = unit(1, "mm"))
  ) +
  
# Color scale and layout adjustments
  scale_color_manual(values = MetBrewer::met.brewer("Klimt")) +
  coord_flip(xlim = c(0.1, 1.5), ylim = c(200, 450), expand = TRUE) +
  guides(color = "none") +
  
# Labels and theme adjustments
  labs(title = "Legend") +
  theme_void(base_family = gp_font_family) +
  theme(
    plot.title = element_text(family = gp_font_family, size = 9, hjust = 0.75, vjust = 1.5),
    plot.background = element_rect(color = "grey30", size = 0.2, fill = gp_color)
  )


```

### Final Plot 
```{r, fig.width=10, fig.height=6, dpi=300}

# Insert the custom legend into the main plot
good_plot_final <- good_plot_main + 
  inset_element(good_plot_legend, l = 0.99, r = 0.7, t = 1, b = 0.5, clip = FALSE)

good_plot_final
```

### Explaination

As a template I used the following template https://r-graph-gallery.com/web-ridgeline-plot-with-inside-plot-and-annotations.html and adjusted it for the chickwts data set from base r.

What Makes This Plot Good:
<ul>
  <li> All essential information is present within the plot, allowing it to be understood independently without additional context. </li>
  <li>Despite containing multiple layers of information, the plot is uncluttered and easy to interpret. </li>
  <li>The custom legend provides a comprehensive and straightforward explanation of the plot elements, helping reader comprehension. </li>
  <li> Visually appealing presentation. </li>
</ul>

## Bad Plot

### Plot Specifications Bad Plot

```{r}

#Import specific font for plot
Rokemon::import_pokefont()

```

### Bad Plot
```{r, warning=FALSE, fig.width=13, fig.height=8, dpi=300}

bad_plot <- data_bad_plot %>% 
  ggplot(aes(y = total_base_stats, x = primary_type)) +
  
# Interval for attack and distribution plots for defence
  stat_interval(aes(y = attack, alpha = 0.8), color = "pink") +
  stat_dist_pointinterval(aes(y = defense, alpha = 0.8), color = "blue") +
  
# Scatter plot for Pokémon with other attributes (category, health point, secondary type)
  geom_point(aes(alpha = 0.8, shape = category, size = hp, color = secondary_type)) +
  
  # Annotate specific Pokémon names
  geom_text(
    aes(label = ifelse(name %in% c("pikachu", "squirtle", "mew", "mewtwo", 
                                   "caterpie", "magikarp", "gyarados", "snorlax"), 
                       name, "")), 
    vjust = -1, size = 3.5
  ) +
  
  # Color scale for secondary types 
  scale_color_poketype() +
  
  # Flip coordinates for better label spacing
  coord_flip() +
  
  # Guide lines 
  geom_hline(yintercept = seq(300, 650, by = 50), color = "gray90", linetype = "dashed") +
  
  # Theme adjustments
  theme_gameboy() +
  theme(
    panel.grid.major.x = element_line(color = "gray90"),  
    plot.title = element_text(face = "bold", size = 20, hjust = 0.5),
    legend.box.margin = margin(t = 10, r = 10, b = 10, l = 10),
    legend.box = "horizontal",
    legend.justification = "left",
    legend.spacing.x = unit(0.5, "cm"),
    legend.spacing.y = unit(0.3, "cm"),
    legend.position = "bottom",
    axis.title = element_text(size = 12),
    axis.text.y = element_text(),
    plot.margin = margin(t = 20, r = 30, b = 30, l = 20)
  ) +
  
# Customized legends
  guides(
    colour = guide_legend(nrow = 3, byrow = TRUE, title.position = "top"),
    size = guide_legend(nrow = 3, title.position = "top"), #Change number of legnend rows 
    shape = guide_legend(nrow = 3, title.position = "top"),
    alpha = guide_legend(nrow = 3, title.position = "top")
  ) +
  
# Annotate intervals
  annotate("text", x = 3, y = 175, label = "Pink: Attack Intervals", color = "pink", size = 4) +
  annotate("text", x = 4, y = 175, label = "Blue: Defense with Distribution", color = "blue", size = 4) +
  
# Labels for plot
  labs(
    title = "Pokemon Stats 1st Generation",
    y = "Total Base Stats",
    x = "Primary Type",
    color = "Secondary Type",
    size = "Health Points",
    shape = "Category",
    alpha = "Mean Defense"
  )

bad_plot
```

### Explaination


What Makes This Plot Bad:
<ul>
<li> It is visually cluttered, hard to interpret, and overplotted. </li>
<li> Attack and defense stats are plotted differently, which makes them seem like they are visualizing the same values with different plots. </li>
<li> Using only one color makes the intervals of the attack stats unreadable. Without an explanation, it’s impossible to understand what precentages these intervals represent. </li>
<li> Total base stats, attack, and defense are all plotted on the same axis. Since the axis is stretched due to the high values of the base stats, it becomes difficult to distinguish differences in attack and defense. Additionally, the x-axis lacks labels in this section of the plot. </li>
<li> The large number of secondary types, along with their color choices and the green background, combined with low transparency, makes it nearly impossible to identify the exact secondary type. </li>
<li> The size of the dots represents the number of health points, but this information is unreadable. The legend shows a circle, but most points are actually squares, which adds to the confusion. </li>
<li> The Pokémon labels are not organized, making it unclear which point is being referenced. Some names are even placed outside the plot. </li>
</ul>

# Session info

```{r}

sessionInfo()

```


